\frame{\frametitle{Course 1} }


\frame{\frametitle{Course 1} \textit{Zaid and Fatimah: the next slide gives the objectives for this course.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementCourse

\begin{itemize}

    \item Characterize a linear systems and sets of vectors using the concepts of span, linear independence, and linear transforms. 
    \item Apply elementary row operations to solve linear systems of equations.
    \item Construct examples of vectors, matrices, and systems that have required properties related to linear transforms, linear independence, and span. 
\end{itemize}

}


\frame{\frametitle{Course 1 Module 1} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Solve linear systems using row reduction operations. 
    \item Describe the structure of a matrix using echelon forms. 
    \item Characterize vectors and systems using the idea of the span of a set of vectors.
\end{itemize}

}

\frame{\frametitle{Course 1 Module 2} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Characterize sets of vectors and solutions to linear systems using the concepts of span, linear combinations, and linear independence. 
\end{itemize}

}

\frame{\frametitle{Course 1 Module 3} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Construct a linear transform to perform a desired mapping. 
    \item Interpret matrix vector multiplication and analyze linear systems using linear transforms. 
\end{itemize}

}
