\frame{\frametitle{Course 2} }


\frame{\frametitle{Course 2} \textit{Zaid and Fatimah: the next slide gives the objectives for this course.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementCourse

\begin{itemize}

    \item Apply the matrix inverse to describe linear systems and their solutions. 
    \item Apply matrix algebra to solve systems efficiently using the LU factorization.
    \item Apply matrix algebra to solve problems in computer graphics and economics.
    \item Apply subspaces to describe the structure of a linear system. 

\end{itemize}

}



\frame{\frametitle{Course 2 Module 1} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Determine when the inverse of a matrix exists.
    \item Calculate the inverse of a matrix when it exists. 
    \item Apply matrix algebra and its properties to analyze and compute matrix products and sums.

\end{itemize}

}

\frame{\frametitle{Course 2 Module 2} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Apply the LU factorization to solve linear systems. 
    \item Apply matrix algebra to solve problems in economics and computer graphics.

    
\end{itemize}

}

\frame{\frametitle{Course 2 Module 3} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Construct a basis for a subspace. 
    \item Determine whether a set of vectors is a subspace. 
    \item Identify the coordinates of a vector in a subspace. 
    
\end{itemize}

}
