\frame{\frametitle{Course 3} }


\frame{\frametitle{Course 3} \textit{Zaid and Fatimah: the next slide gives the objectives for this course.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Use determinants to characterize linear transforms and the invertibility of a matrix.
    \item Apply eigenvalue theorems to analyze systems and matrices.
    \item Apply eigenvalues and eigenvectors to study Markov Chains, and the Google Matrix algorithm.

\end{itemize}

}


\frame{\frametitle{Course 2 Module 1} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Apply theorems related to a determinant to compute determinants efficiently.
    \item Identify whether a matrix is invertible by using a determinant.
    \item Describe how areas change under linear transforms using determinants.

\end{itemize}

}

\frame{\frametitle{Course 2 Module 2} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Solve Markov chain problems to determine steady-state vectors.
    \item Calculate the eigenvectors and eigenvalues of a matrix. 
    \item Apply theorems related to eigenvalues and eigenvectors to characterize a matrix and systems of equations. 
    
\end{itemize}

}

\frame{\frametitle{Course 2 Module 3} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Apply the google page rank algorithm to model how users are navigating the web. 
    \item Calculate eigenvalues and eigenvectors that are complex. 
    \item Construct a matrix factorization that uses complex eigenvalues and eigenvectors to represent a matrix as a product of real matrices. 
    
\end{itemize}

}
