\frame{\frametitle{Course 4} }


\frame{\frametitle{Course 4} \textit{Zaid and Fatimah: the next slide gives the objectives for this course.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementCourse

\begin{itemize}

    \item Apply theorems related to orthogonality to construct mathematical models for data using least-squares.
    \item Apply theorems related to orthogonal complements, and their relationships to subspaces, to characterize vectors and linear systems.
    \item Apply eigenvalues and eigenvectors to solve optimization problems that are subject to distance and orthogonality constraints.
    \item Apply the singular value decomposition to characterize the structure of a matrix and its invertibility.
\end{itemize}

}




\frame{\frametitle{Course 4 Module 1} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }


\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Project vectors onto subspaces using orthogonality.
    \item Characterize matrices in terms of the four fundamental subspaces.
    \item Describe relationships between vectors using dot products and orthogonality.
\end{itemize}

}



\frame{\frametitle{Course 4 Module 2} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Construct an orthogonal basis for a subspace given any basis for that space.
    \item Use the QR factorization to decompose a matrix into the product of two matrices, one of which gives an orthonormal basis for the column space of a matrix. 
    
\end{itemize}

}

\frame{\frametitle{Course 4 Module 3} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Apply the least-squares algorithm to approximate solutions to inconsistent systems.
    \item Apply the least squares algorithm to a range of data science problems.
    \item Characterize how well the least-squares algorithm fit a given data set using orthogonal projections.

\end{itemize}

}

\frame{\frametitle{Course 4 Module 4} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Construct an orthogonal diagonalization of a symmetric matrix.
    \item Represent and analyze quadratic forms using symmetric matrices. 
    \item Apply symmetric matrices to solve constrained optimization problems. 

\end{itemize}

}

\frame{\frametitle{Course 4 Module 5} \textit{Zaid and Fatimah: the next slide gives the objectives for this module.}  }

\frame{\frametitle{Learning Objectives}

\LearningObjectiveStatementModule

\begin{itemize}

    \item Construct the SVD of a matrix. 
    \item Apply the SVD to determine the rank of a matrix, construct bases for the four fundamental subspaces, and estimate the condition number of a matrix. 
\end{itemize}

}