\title{Matrices with Orthonormal Columns}
\subtitle{\SubTitleName}
\institute[]{\Course}
\author{\Instructor}
\maketitle   
  

 
\begin{frame}\frametitle{Topics and Objectives}
\Emph{Topics} \\
%\TopicStatement
\begin{itemize}

    \item matrices with orthonormal columns and their properties
    

\end{itemize}

\vspace{0.5cm}

\Emph{Learning Objectives}\\

\LearningObjectiveStatement

\begin{itemize}
    \item apply the properties of matrices with orthonormal columns to determine whether mathematical statements that involve them are accurate
    
    % \item express a vector as a linear combination of orthogonal vectors
    % \item characterize bases for subspaces of $\mathbb R^n$, and

\end{itemize}

\end{frame}






\begin{frame}{Orthonormal Columns}

    \begin{center}\begin{tikzpicture} \node [mybox](box){\begin{minipage}{0.95\textwidth}\vspace{4pt}
        An $ m \times n$ matrix $ U$ has orthonormal columns if and only if $ U ^{T} U = I_n $. 
    \end{minipage}};
    \node[fancytitle, right=10pt] at (box.north west) {Theorem};
    \end{tikzpicture}\end{center}

    \pause
    Can $U$ have orthonormal columns if $n > m$? 

\end{frame}





\begin{frame}{Properties of Matrices with Orthonormal Columns }
\begin{center}\begin{tikzpicture} \node [mybox](box){\begin{minipage}{0.85\textwidth}\vspace{4pt}
    Suppose  $ m \times n$ matrix $U$ has orthonormal columns and $\vec x$ and $\vec y$ are vectors in $\mathbb R^n$. 

    \begin{enumerate}
        \item<2-> $ \lVert U \vec x\rVert = \lVert \vec x \rVert$
        \item<3-> $ (U \vec x  ) \cdot (U \vec y) = \vec x \cdot \vec y$
        \item<4-> $ (U \vec x  ) \cdot (U \vec y) = 0$ if and only if $\vec x \cdot \vec y = 0$
    \end{enumerate}

    \end{minipage}};
    \node[fancytitle, right=10pt] at (box.north west) {Theorem};
    \end{tikzpicture}\end{center}
    
    \pause
    \onslide<5->{These properties tell us that the mapping $\vec x \to U\vec x$ preserves lengths and orthogonality.} 
    \onslide<6->{Proofs of the above statements are similar. Let's prove the first statement.}

\end{frame}


\begin{frame}{Proof That $ \lVert U \vec x\rVert = \lVert \vec x \rVert$}
    
    First consider $\lVert U \vec x\rVert^2$. 
    \begin{align*}
        \onslide<2->{ \lVert U \vec x\rVert^2 &= (U \vec x ) \cdot (U \vec x)} 
        \onslide<4->{  =  \vec x^T U^T U \vec x } 
    \end{align*}
    \onslide<5->{But $U^TU = I$.}
    \begin{align*}
        \onslide<5->{ \lVert U \vec x\rVert^2 &= \vec x^T I \vec x } 
        \onslide<6->{  =  \lVert \vec x\rVert^2 } 
    \end{align*}
    \onslide<7->{Taking the square root of both sides yields $$\lVert U \vec x\rVert = \lVert \vec x \rVert$$}
    
\end{frame}






\begin{frame}{Orthogonal Matrices}

    \begin{itemize}
        \item<2-> An \Emph{orthogonal matrix} is a square matrix whose columns are orthonormal.
        \item<3-> A better name might be \textit{orthonormal matrix}, but \textit{orthogonal matrix} is a standard term in linear algebra. 
        \item<4-> Also, matrices with columns that are orthogonal, but do not have unit length, are not often used. 
        \item<5-> Note that if $U$ is an orthogonal matrix, then $U^{-1} = U^T$. 
    \end{itemize}


\end{frame}

\begin{frame}{Determinant of an Orthogonal Matrix}

    If $A$ is a square matrix with orthonormal columns, then $\det A$ is equal to $1$ or to $-1$. 

    \vspace{12pt}
    
    \onslide<2->{\Emph{Proof}}
    \begin{align*}
        \onslide<3->{1 = \det I = } 
        \onslide<4->{\det(A^TA) =} 
        \onslide<5->{\det(A^T)\det(A) =}
        \onslide<6->{(\det A)^2}
    \end{align*}
    
\end{frame}


\begin{frame}{Examples: Orthogonal Matrices}  

    Indicate whether the statements are true or false.
    \begin{enumerate}

    \item If $U$ is orthogonal, then its columns are also linearly independent. 

    \item If the determinant of a square matrix is 1, then the matrix must be orthogonal.  
    
    \end{enumerate}
\end{frame}

\frame{\frametitle{Summary}

    \SummaryLine \vspace{4pt}
    \begin{itemize}\setlength{\itemsep}{8pt}

    \item matrices with orthonormal columns and their properties
    
    \end{itemize}
    
    \vspace{8pt}
    
    \onslide<2->{Properties:}
    \begin{enumerate}
        \item<2-> $ \lVert U \vec x\rVert = \lVert \vec x \rVert$
        \item<2-> $ (U \vec x  ) \cdot (U \vec y) = \vec x \cdot \vec y$
        \item<2-> $ (U \vec x  ) \cdot (U \vec y) = 0$ if and only if $\vec x \cdot \vec y = 0$
    \end{enumerate}
    
    \vspace{8pt}
    
    \onslide<3->{If $U$ is square and has orthonormal columns, then
    \begin{itemize}
        \item $U$ is orthogonal
        \item $\det U$ is equal to 1 or -1
    \end{itemize}
    }
    \pause
    
}

\frame{\frametitle{Blank}
}


